---
title: "Zeyd Khalil, HW10, October 29, 2020"
output: pdf_document
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)

dat <-
  read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv")
deaths_by_state <- dat %>%
  group_by(state,date) %>%
  summarize(total_deaths=sum(deaths)) %>%
  ungroup() %>%
  filter(date == max(date)) %>%
  arrange(desc(total_deaths))
```

# Exercise 1 - create a tibble that has both deaths and total cases per state, arranged by the total number of deaths in descending order   

```{r}
state.1 <- 
  dat %>% 
  group_by(state, date) %>% 
  summarize(total_deaths = sum(deaths), total_cases = sum(cases))

state.1

deaths <- 
  state.1 %>% 
  group_by(state) %>% 
  filter(total_deaths == max(total_deaths), total_cases == max(total_cases)) %>% 
  arrange(desc(total_deaths)) %>% 
  distinct() 

deaths
```

# Exercise 2 - 

```{r, warning=FALSE}
dat_dmv <- dat %>% 
  filter(state == "District of Columbia" | state == "Virginia" | state == "Maryland", county == "Anne Arundel" | county == "Charles" | county == "Frederick" | county == "Howard" | county == "Montgomery" | county == "Prince George's" | county == "Alexandria City" | county == "Arlington" | county == "Fairfax" | county == "Loudoun" | county == "Prince William" | county == "District of Columbia")

dat_dmv %>% 
   ggplot(aes(x = date, y = cases, group = county, col = county)) +
   geom_line() +
   geom_point() +
   facet_wrap(~ state) +
   scale_y_log10() +
   scale_color_brewer(palette = "Set1")
``` 

# Exercise 3 - 

```{r}
dmv_total_cases <- 
  dat_dmv %>% 
  group_by(date) %>% 
  summarize(total_cases = sum(cases))  

dmv_total_cases %>% 
  ggplot(aes(x = date, y = total_cases)) + 
  geom_point() 

dmv_total_deaths <- 
  dat_dmv %>% 
  group_by(date) %>% 
  summarise(total_deaths = sum(deaths))  

dmv_total_deaths %>% 
  ggplot(aes(x = date, y = total_deaths)) + 
  geom_point()
```  

> Ask TA About this tomorrow 

# Exercise 4 - Read in and tidy both the global and US datasets. For the US data, produce a second tidy dataset called US_by_state that has the total of Confirmed cases, deaths and population for each date for each state.  

```{r}
cases_global <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")

deaths_global <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")

cases_us <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv")

deaths_us <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv")
```  

> Tidying the datasets

```{r}
cases_global <- 
  cases_global %>% 
  pivot_longer(cols = c(`1/22/20`:`10/28/20`), names_to = "date", values_to = "cases")

deaths_global <- 
  deaths_global %>% 
  pivot_longer(cols = c(`1/22/20`:`10/28/20`), names_to = "date", values_to = "Deaths")

cases_us <- 
  cases_us %>% 
  pivot_longer(cols = c(`1/22/20`:`10/28/20`), names_to = "date", values_to = "cases")

deaths_us <- 
  deaths_us %>% 
  pivot_longer(cols = c(`1/22/20`:`10/28/20`), names_to = "date", values_to = "Deaths")

```


```{r}
global <- cases_global %>% full_join(deaths_global) %>%
  rename(Country_Region = `Country/Region`, Province_State = `Province/State`) 

global
```
```{r}
US <- deaths_us %>%
  full_join(cases_us, by = c("Combined_Key", "date", "Admin2", "Province_State", "Country_Region")) %>%
    rename(Long = Long_.x, Lat = Lat.x)  %>%
    select(Admin2, Province_State, Country_Region, Lat, Long, Population, date, cases, Deaths)

```

```{r}
US_by_state <- US %>% group_by(Province_State, Country_Region, date) %>% summarize(cases = sum(cases), Deaths = sum(Deaths), Lat = median(Lat), Long = median(Long), Population = sum(Population)) %>% mutate(Deaths_per_mill = (Deaths / Population) * 1000000) %>% select(Province_State, Country_Region, date, cases, Deaths, Deaths_per_mill, Population, Lat, Long) %>% ungroup() 

US_by_state
```  

# Exercise 5 - Replace the US observations in the global dataset with the US data. Add a new variable called continent to the dataset. Be sure there are no NA’s for continent. Also create a new variable Country_State that comines the Province_State with Country_Region.

```{r expand_US, echo=FALSE}
exp_global <- 
  global %>% 
  filter(Country_Region != "US")

global_by_country <- 
  full_join(US_by_state, exp_global)
```

# Exercise 7 - 

```{r}
top_25 <- cases_global %>% select(`Country/Region`, cases) %>% group_by(`Country/Region`) %>% summarize(Confirmed_cases = max(cases)) %>% arrange(desc(Confirmed_cases)) %>% head(n = 25)
```  

